Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-08 evaluation.tex
Formatting changed as per PACM rules.

  let deregister_student_txn sid = 
    let s_enrs = SQL.select [Enrollment] (fun e -> e.e_s_id = sid) in
    if Set.is_empty s_enrs then
      SQL.delete Student (fun s -> s.s_id = sid)
    else ()
\end{ocaml}
\caption{Courseware Application}
\label{fig:courseware_code}
\end{figure}

\textbf{Courseware} The Courseware application allows new courses to be
added (via an \C{add\_course} transaction), and new students to be
registered (via a \C{register\_student} transaction) into a database. A registered
student can enroll (\C{enroll}) in an existing course,
provided that enrollment has not already exceeded the course
capacity (\C{c\_capacity}). A course with no enrollments can be
canceled (\C{cancel\_course}). Likewise, a student who is not enrolled
in any course can be deregistered (\C{deregister\_student}). Besides
\C{Student} and \C{Course} tables, there is also an \C{Enrollment}
table to track the many-to-many enrollment relationship between
courses and students. The simplified code for the Courseware
application with only \C{enroll} and \C{deregister\_student}
transactions is shown in Fig.~\ref{fig:courseware_code}. The
application is required to preserve the following invariants on the
database:

\begin{enumerate}
\item  $I_1$: An enrollment record should always refer to an existing student and an existing course.
\item  $I_2$: The capacity (\C{c\_capacity}) of a course should always be a
  non-negative quantity.
\end{enumerate}
\noindent Both $I_1$ and $I_2$ can be violated under weak isolation.
$I_1$ can be violated, for example, when \C{deregister\_student} runs
concurrently with \C{enroll}, both at RC isolation. While the former
transaction removes the student record after checking that no
enrollments for that student exists, the latter transaction
concurrently adds an enrollment record after checking the student
exists.  Both can succeed concurrently, resulting in an invalid state.
Invariant $I_2$ can be violated by two \C{enroll}s, both reading
\C{c\_capacity}=1, and both (atomically) decrementing it, resulting in
\C{c\_capacity}=-1.  We ran \thetool on the Courseware application
(Fig.~\ref{fig:courseware_code}) after annotating transactions with
their respective guarantees, and asserting $I = I_1 \wedge I_2$ as the
correctness condition. The guarantees $G_e$ and $G_d$ for \C{enroll}
and \C{deregister\_student} transactions, respectively, are shown below:
\begin{smathpar}
  \begin{array}{lcl}
    G_e(\stg,\stg') & \Leftrightarrow & \stg_s'=\stg_s
      \conj \exists\C{cid}.\exists\C{sid}.\\
    & & \hspace*{0.3in} \stg_c' = \stg_c \bind \lambda c.~
        \itel{c.\C{c\_id}=\C{cid}\\
    & & \hspace*{1.15in}}
          {\existsl(c',~c'.\C{id}=c.\C{id} \wedge
              c'.\C{c\_name}=c.\C{c\_name} \wedge c'.\C{c\_capacity}\ge0,
              ~\{c'\})\\
    & & \hspace*{1.15in}}
          {\{c\}}\\
    & & \hspace*{0.15in}\conj \stg_e = \stg_e' \bind \lambda e.~ 
        \itel{e.\C{e\_c\_id}=\C{cid} \wedge e.\C{e\_s\_id}=\C{sid}}
          {\emptyset}{\{e\}}\\
%   & & \hspace*{1.15in}} 
%         {\emptyset} {\{e\}}\\
    G_d(\stg,\stg') & \Leftrightarrow & \stg_c' = \stg_c \conj
      \stg_e' = \stg_e \conj \exists \C{sid}.~
      \itel{\forall(e \in \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
%   & & \hspace*{0.3in} \itel{\forall(e \in
%           \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s \bind \lambda s.~
           \itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s}\\
  \end{array}
\end{smathpar}
For the sake of this presentation we split $\stg$ into three disjoint
sets of records, $\stg_s$, $\stg_c$, and $\stg_e$, standing for
\C{Student}, \C{Course}, and \C{Enrollment} tables, respectively.
Observing that the set language $\SL$ (Sec.~\ref{sec:inference}),
besides being useful for automatic verification, also facilitates
succinct expression of transaction semantics, we define $G_e$ and
$G_d$ in a combination of FOL and $\SL$. $G_e$ essentially says that
the \C{enroll} transaction leaves the \C{Student} table unchanged,
while it may update the \C{c\_capacity} field of a \C{Course} record
to a non-negative value (even when it doesn't update, it is the case
that $c'.\C{c\_capacity}\ge0$, because $c'=c$, and $c\in\stg_c$, and
we know that $I_2(\stg_c)$). $G_e$ also conveys that \C{enroll} might
insert a new \C{Enrollment} record by stating that $\stg_e$, the
\C{Enrollment} table in the pre-state, contains all records $e$ from
$\stg_e'$, the table in the post-state, except when $e.\C{e\_c\_id}$ and
$e.\C{e\_s\_id}$ match \C{cid} and \C{sid}, respectively. The
guarantee $G_d$ of \C{deregister\_student} asserts that the
transaction doesn't write to \C{Course} and \C{Enrollment} tables. The
transaction might however delete a \C{Student} record bearing an
\C{id}=\C{sid} (formally, $\stg_s' = \stg_s \bind \lambda s.~
\itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}$), for some \C{sid} for
which no corresponding \C{Enrollment} records are present in the
pre-state (in other words, $\forall(e \in
\stg_e).~e.\C{e\_s\_id}\neq\C{sid}$).

With help of the guarantees, such as those described above, \thetool
was able to automatically discover the aforementioned anomalous
executions, and was subsequently able to infer that the anomalies can
be preempted by promoting the isolation level of \C{enroll} and
\C{deregister\_student} to SER (on both MySQL and PostgreSQL), leaving
the isolation levels of remaining transactions at RC. The total time
for inference and verification took less than a minute running on a
conventional laptop.

>>>>

  let deregister_txn sid = 
    let s_enrs = SQL.select [Enrollment] (fun e -> e.e_s_id = sid) in
    if Set.is_empty s_enrs then
      SQL.delete Student (fun s -> s.s_id = sid)
    else ()
\end{ocaml}
\caption{Courseware Application}
\label{fig:courseware_code}
\end{figure}

\textbf{Courseware} The Courseware application allows new courses to be
added (via an \C{add\_course} transaction), and new students to be
registered (via a \C{register} transaction) into a database. A registered
student can enroll (\C{enroll}) in an existing course,
provided that enrollment has not already exceeded the course
capacity (\C{c\_capacity}). A course with no enrollments can be
canceled (\C{cancel\_course}). Likewise, a student who is not enrolled
in any course can be deregistered (\C{deregister}). Besides
\C{Student} and \C{Course} tables, there is also an \C{Enrollment}
table to track the many-to-many enrollment relationship between
courses and students. The simplified code for the Courseware
application with only \C{enroll} and \C{deregister}
transactions is shown in Fig.~\ref{fig:courseware_code}. The
application is required to preserve the following invariants on the
database:

\begin{enumerate}
\item  $I_1$: An enrollment record should always refer to an existing student and an existing course.
\item  $I_2$: The capacity (\C{c\_capacity}) of a course should always be a
  non-negative quantity.
\end{enumerate}
\noindent Both $I_1$ and $I_2$ can be violated under weak isolation.
$I_1$ can be violated, for example, when \C{deregister} runs
concurrently with \C{enroll}, both at RC isolation. While the former
transaction removes the student record after checking that no
enrollments for that student exists, the latter transaction
concurrently adds an enrollment record after checking the student
exists.  Both can succeed concurrently, resulting in an invalid state.
Invariant $I_2$ can be violated by two \C{enroll}s, both reading
\C{c\_capacity}=1, and both (atomically) decrementing it, resulting in
\C{c\_capacity}=-1.  We ran \thetool on the Courseware application
(Fig.~\ref{fig:courseware_code}) after annotating transactions with
their respective guarantees, and asserting $I = I_1 \wedge I_2$ as the
correctness condition. The guarantees $G_e$ and $G_d$ for \C{enroll}
and \C{deregister} transactions, respectively, are shown below:
\begin{smathpar}
  \begin{array}{lcl}
    G_e(\stg,\stg') & \Leftrightarrow & \stg_s'=\stg_s
      \conj \exists\C{cid}.\exists\C{sid}.\\
    & & \hspace*{0.3in} \stg_c' = \stg_c \bind \lambda c.~
        \itel{c.\C{c\_id}=\C{cid}\\
    & & \hspace*{1.15in}}
          {\existsl(c',~c'.\C{id}=c.\C{id} \wedge
              c'.\C{c\_name}=c.\C{c\_name} \\
    & & \hspace*{2.6in}\wedge~ c'.\C{c\_capacity}\ge0, ~\{c'\})\\
    & & \hspace*{1.15in}}
          {\{c\}}\\
    & & \hspace*{0.15in}\conj \stg_e = \stg_e' \bind \lambda e.~ 
        \itel{e.\C{e\_c\_id}=\C{cid} \wedge e.\C{e\_s\_id}=\C{sid}}
          {\emptyset}{\{e\}}\\
%   & & \hspace*{1.15in}} 
%         {\emptyset} {\{e\}}\\
    G_d(\stg,\stg') & \Leftrightarrow & \stg_c' = \stg_c \conj
      \stg_e' = \stg_e \conj \exists \C{sid}.~
      \itel{\forall(e \in \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
%   & & \hspace*{0.3in} \itel{\forall(e \in
%           \stg_e).~e.\C{e\_s\_id}\neq\C{sid}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s \bind \lambda s.~
           \itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}\\
    & & \hspace*{1.55in}}
          {\stg_s' = \stg_s}\\
  \end{array}
\end{smathpar}
For the sake of this presentation we split $\stg$ into three disjoint
sets of records, $\stg_s$, $\stg_c$, and $\stg_e$, standing for
\C{Student}, \C{Course}, and \C{Enrollment} tables, respectively.
Observing that the set language $\SL$ (Sec.~\ref{sec:inference}),
besides being useful for automatic verification, also facilitates
succinct expression of transaction semantics, we define $G_e$ and
$G_d$ in a combination of FOL and $\SL$. $G_e$ essentially says that
the \C{enroll} transaction leaves the \C{Student} table unchanged,
while it may update the \C{c\_capacity} field of a \C{Course} record
to a non-negative value (even when it doesn't update, it is the case
that $c'.\C{c\_capacity}\ge0$, because $c'=c$, and $c\in\stg_c$, and
we know that $I_2(\stg_c)$). $G_e$ also conveys that \C{enroll} might
insert a new \C{Enrollment} record by stating that $\stg_e$, the
\C{Enrollment} table in the pre-state, contains all records $e$ from
$\stg_e'$, the table in the post-state, except when $e.\C{e\_c\_id}$ and
$e.\C{e\_s\_id}$ match \C{cid} and \C{sid}, respectively. The
guarantee $G_d$ of \C{deregister} asserts that the
transaction doesn't write to \C{Course} and \C{Enrollment} tables. The
transaction might however delete a \C{Student} record bearing an
\C{id}=\C{sid} (formally, $\stg_s' = \stg_s \bind \lambda s.~
\itel{s.\C{id}=\C{sid}}{\emptyset}{\{s\}}$), for some \C{sid} for
which no corresponding \C{Enrollment} records are present in the
pre-state (in other words, $\forall(e \in
\stg_e).~e.\C{e\_s\_id}\neq\C{sid}$).

With help of the guarantees, such as those described above, \thetool
was able to automatically discover the aforementioned anomalous
executions, and was subsequently able to infer that the anomalies can
be preempted by promoting the isolation level of \C{enroll} and
\C{deregister} to SER (on both MySQL and PostgreSQL), leaving
the isolation levels of remaining transactions at RC. The total time
for inference and verification took less than a minute running on a
conventional laptop.

### 2017-11-08 implementation.tex
Formatting changed as per PACM rules.

We have implemented our DSL to define transactions as monadic
computations in OCaml (modulo some syntactic sugar), and our automatic
reasoning framework as an extra frontend pass (called \thetool) in the
ocamlc 4.03 compiler\footnote{The source code is available at
available at \emph{https://goo.gl/nmUXZK}}. The input to \thetool is a
program in our DSL that describes the schema of the database as a
collection of OCaml type definitions, and transactions as OCaml
functions, whose top-level expression is an application of the
\C{atomically\_do} combinator. For instance, TPC-C's schema from
Fig.~\ref{fig:schema} can be described via the OCaml type definitions
shown in Fig.~\ref{fig:ocaml-schema}.  \thetool also requires a
specification of the program in the form of a collection of guarantees
($G$), one per transaction, and an invariant $I$ that is a conjunction
of the integrity constraints on the database. An auxiliary DSL that
includes the first-order logic (FOL) combinators has been implemented
for this purpose. \thetool's verification pass follows OCaml's type
checking pass, hence the concrete artifact of verification is OCaml's
typed AST. The tool is already equipped with  an axiomatization of
PostgreSQL and MySQL's isolation levels expressed in our FOL DSL.
Other data stores can be similarly axiomatized. The concrete result of
verification is an assignment of an isolation level of the selected
data store to each transaction in the program.

>>>>

We have implemented our DSL to define transactions as monadic
computations in OCaml (modulo some syntactic sugar), and our automatic
reasoning framework as an extra frontend pass (called \thetool) in the
ocamlc 4.03 compiler\footnote{The source code is available at
available at \url{https://github.com/gowthamk/acidifier}}. The input
to \thetool is a program in our DSL that describes the schema of the
database as a collection of OCaml type definitions, and transactions
as OCaml functions, whose top-level expression is an application of
the \C{atomically\_do} combinator. For instance, TPC-C's schema from
Fig.~\ref{fig:schema} can be described via the OCaml type definitions
shown in Fig.~\ref{fig:ocaml-schema}.  \thetool also requires a
specification of the program in the form of a collection of guarantees
($G$), one per transaction, and an invariant $I$ that is a conjunction
of the integrity constraints on the database. An auxiliary DSL that
includes the first-order logic (FOL) combinators has been implemented
for this purpose. \thetool's verification pass follows OCaml's type
checking pass, hence the concrete artifact of verification is OCaml's
typed AST. The tool is already equipped with  an axiomatization of
PostgreSQL and MySQL's isolation levels expressed in our FOL DSL.
Other data stores can be similarly axiomatized. The concrete result of
verification is an assignment of an isolation level of the selected
data store to each transaction in the program.

### 2017-11-08 inference.tex
edits
Formatting changed as per PACM rules.

Sec.~\ref{sec:motivation}, which is a function $\T$ of form
$\lambda(\stl,\stg).~\stl \cup \F(\stg)$. Nonetheless, we keep
referring to $\F$ as state transformer. Since the execution is subject

@@

    \F_I & = & \lambda(\stg).~ \{\langle\C{ol\_o\_id}=\C{k2},\,
                 \C{ol\_d\_id}=\C{k3},\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                 \C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}\\

@@

    \phi_U & = & \hspace*{0.2in}\forall \C{item\_req}.\forall s. \forall s'.~ 

@@

    & & \conj \hspace*{0.08in}\forall \C{item\_req}.\forall s'. \exists s.~ 

@@

    & & \hspace*{0.28in} g_0(\C{item\_req},s') \Rightarrow (s \in \stg) \conj

@@

    & & \hspace*{1.8in}}

@@

    & & \hspace*{1.8in}} {\bot}) \\

@@

        r = \langle \C{ol\_o\_id}=\C{k2},\,
                    \C{ol\_d\_id}=\C{k3},\, 
                    \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                    \C{ol\_qty}=\C{item\_req.ol\_qty}
             \rangle\\

@@

$(\phi,\G)$, as defined below:

@@

    \phi & = & \phi_I \conj \phi_U \conj \\
    & & \hspace*{0.2in}\forall \C{item\_req}. \forall s.~ 
        \pi_3 \Leftrightarrow (\C{item\_req} \in \C{item\_reqs} \conj 
        (\G_U(\C{item\_req},s') \vee \G_I(\C{item\_req},s'))
        \Rightarrow g_1(s))\\
    & & \conj \hspace*{0.05in}\forall s.\exists \C{item\_req}. ~ 
        \pi_4 \Leftrightarrow (g_1(s) \Rightarrow (\C{item\_req} \in \C{item\_reqs} 
        \conj (\G_U(\C{item\_req},s') \vee \G_I(\C{item\_req},s')))\\

>>>>

Sec.~\ref{sec:motivation}, which is a function $\T$ of the form
$\lambda(\stl,\stg).~\stl \cup \F(\stg)$. Nonetheless, for simplicity,
we will continue to refer to $\F$ as state transformer. Since the execution is subject

@@

    \F_I & = & \lambda(\stg).~ \{\langle\C{ol\_o\_id}=\C{k2};\,
                 \C{ol\_d\_id}=\C{k3};\,
                 \C{ol\_i\_id}=\C{item\_req.ol\_i\_id};\,\\
         &   & \hspace*{2.3in}\C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}\\

@@

    \phi_U & = & \hspace*{0.1in}\forall \C{item\_req}.\forall s. \forall s'.~ 

@@

    & & \hspace*{-0.05in}\conj \forall \C{item\_req}.\forall s'. \exists s.~ 

@@

    & & \hspace*{0.15in} g_0(\C{item\_req},s') \Rightarrow s \in \stg \conj
        \itec{{\sf table}(s)=\C{Stock} \wedge 

@@

    & & \hspace*{1.6in}}

@@

    & & \hspace*{1.6in}} {\bot} \\

@@

        r = \langle \C{ol\_o\_id}=\C{k2};\,
                    \C{ol\_d\_id}=\C{k3};\, 
                    \C{ol\_i\_id}=\C{item\_req.ol\_i\_id};\,\\
    & & \hspace*{3in}\C{ol\_qty}=\C{item\_req.ol\_qty} \rangle\\

@@

$(\phi_I \wedge \phi_U \wedge \phi_1 \wedge \phi_2,\G)$, where
$\phi_1$, $\phi_2$, and $\G$ are as defined below:

@@

    \phi_1 & = & \forall \C{item\_req}. \forall s.~ 
        \pi_3 \Leftrightarrow \C{item\_req} \in \C{item\_reqs} \wedge
        \G_U(\C{item\_req},s') \vee \G_I(\C{item\_req},s')
        \Rightarrow g_1(s)\\
    \phi_2 & = & \forall s.\exists \C{item\_req}. ~ 
        \pi_4 \Leftrightarrow g_1(s) \Rightarrow \C{item\_req} \in \C{item\_reqs} 
        \wedge \G_U(\C{item\_req},s') \vee \G_I(\C{item\_req},s')\\

### 2017-11-08 motivation.tex
Formatting changed as per PACM rules.

\begin{figure}[t]
  \centering
	\begin{subfigure}{0.4\textwidth}
	  \includegraphics[width=0.9\textwidth]{Figures/schema1}
    \caption{A valid TPC-C database. The only existing order belongs
      to the district with \C{d\_id}=11. Its id (\C{o\_id}) is one
      less than the district's \C{d\_next\_o\_id}, and its order
      count (\C{o\_ol\_cnt}) is equal to the number of order line
      records whose \C{ol\_o\_id} is equal to the order's id.  }
		\label{fig:tpcc_db1}
	\end{subfigure}
  \hspace*{.4in}
	\begin{subfigure}{0.52\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/schema2}
    \caption{The database in Fig.~\ref{fig:tpcc_db1} after correctly
      executing a \C{new\_order} transaction. A new order record is
      added whose \C{o\_id} is equal to the \C{d\_next\_o\_id} from
      Fig.~\ref{fig:tpcc_db1}. The district's \C{d\_next\_o\_id} is
      incremented. The order's \C{o\_ol\_cnt} is 2, reflecting the
      actual number of order line records whose \C{ol\_o\_id} is equal
      to the order's id (2).}
		\label{fig:tpcc_db2}
   	\end{subfigure}
\caption{Database schema of TPC-C's order management system.
  The naming
  convention indicates primary keys and foreign keys. For e.g.,
  \C{ol\_id} is the primary key column of the order line table,
  whereas \C{ol\_o\_id} is a foreign key that refers to the \C{o\_id}
  column of the order table.}
\label{fig:schema}
\end{figure}

@@

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  Under weak
isolation\footnote{Weak isolation does not violate atomicity as long
  as the witnessed effects are those of committed transactions},
however, interferences of various kinds are permitted, leading to
executions superficially similar to executions permitted by concurrent
(racy) programs~\cite{GHE15,HPQ+15}.  To illustrate, consider the
behavior of the \C{new\_order} transaction when executed under a
\emph{Read Committed} (RC) isolation level, the default isolation
level in 8 of the 18 databases studied in~\cite{bailishotos}.
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
An executing RC transaction is isolated from \emph{dirty writes},
i.e., writes of uncommitted transactions, but is allowed to witness
the writes of concurrent transactions as soon as they are committed.
Thus, with two concurrent instances of the \C{new\_order} transaction
(call them $T_1$ and $T_2$), both concurrently placing new orders for
different customers under the same district (\C{d\_id}), RC isolation
allows the execution shown in Fig.~\ref{fig:new_order_exec}.

@@

The transformer ($\T_I(\delta,\Delta)$) for the subsequent \C{insert}
statement can be similarly constructed:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup
       \{\langle\C{ol\_o\_id}=\C{k2},\,
                 \C{ol\_d\_id}=\C{k3},\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id},\, 
                 \C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}

>>>>

\begin{figure}[t]
  \centering
	\begin{subfigure}{0.4\textwidth}
	  \includegraphics[width=0.9\textwidth]{Figures/schema1}
    \caption{A valid TPC-C database. The only existing order belongs
      to the district with \C{d\_id}=11. Its id (\C{o\_id}) is one
      less than the district's \C{d\_next\_o\_id}, and its order
      count (\C{o\_ol\_cnt}) is equal to the number of order line
      records whose \C{ol\_o\_id} is equal to the order's id.  }
		\label{fig:tpcc_db1}
	\end{subfigure}
  \hspace*{.2in}
	\begin{subfigure}{0.52\textwidth}
		\includegraphics[width=0.9\textwidth]{Figures/schema2}
    \caption{The database in Fig.~\ref{fig:tpcc_db1} after correctly
      executing a \C{new\_order} transaction. A new order record is
      added whose \C{o\_id} is equal to the \C{d\_next\_o\_id} from
      Fig.~\ref{fig:tpcc_db1}. The district's \C{d\_next\_o\_id} is
      incremented. The order's \C{o\_ol\_cnt} is 2, reflecting the
      actual number of order line records whose \C{ol\_o\_id} is equal
      to the order's id (2).}
		\label{fig:tpcc_db2}
   	\end{subfigure}
\caption{Database schema of TPC-C's order management system.
  The naming
  convention indicates primary keys and foreign keys. For e.g.,
  \C{ol\_id} is the primary key column of the order line table,
  whereas \C{ol\_o\_id} is a foreign key that refers to the \C{o\_id}
  column of the order table.}
\label{fig:schema}
\end{figure}

@@

\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  Under weak
isolation\footnote{Weak isolation does not violate atomicity as long
  as the witnessed effects are those of committed transactions},
however, interferences of various kinds are permitted, leading to
executions superficially similar to executions permitted by concurrent
(racy) programs~\cite{GHE15,HPQ+15}.  To illustrate, consider the
behavior of the \C{new\_order} transaction when executed under a
\emph{Read Committed} (RC) isolation level, the default isolation
level in 8 of the 18 databases studied in~\cite{bailishotos}.
An executing RC transaction is isolated from \emph{dirty writes},
i.e., writes of uncommitted transactions, but is allowed to witness
the writes of concurrent transactions as soon as they are committed.
Thus, with two concurrent instances of the \C{new\_order} transaction
(call them $T_1$ and $T_2$), both concurrently placing new orders for
different customers under the same district (\C{d\_id}), RC isolation
allows the execution shown in Fig.~\ref{fig:new_order_exec}.

@@

The transformer ($\T_I(\delta,\Delta)$) for the subsequent \C{insert}
statement can be similarly constructed:
\begin{smathpar}
\begin{array}{l}
  \lambda(\stl,\stg).~ \stl \cup
       \{\langle\C{ol\_o\_id}=\C{k2};\,
                 \C{ol\_d\_id}=\C{k3};\, \C{ol\_i\_id}=\C{item\_req.ol\_i\_id};\, 
                 \C{ol\_qty}=\C{item\_req.ol\_qty}\rangle\}
